---
output: 
  stevetemplates::article:
    fig_caption: true
#bibliography: master.bib
title: "Estimating process-based model parameters from species distribution data"
subtitle: "Supplementary Appendix G"
author:
- name: Victor Van der Meersch
  affiliation: CEFE, CNRS
- name: Isabelle Chuine
  affiliation: CEFE, CNRS
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: libertine
fontsize: 11pt
# spacing: double
appendix: yes
appendixletter: G
endnote: no
header-includes:
  - \usepackage{pdflscape}
  - \usepackage{xcolor}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage[width=.9\textwidth, textfont=it, font=small]{caption}
  - \usepackage{libertine}

---

```{r setup_appendixG, include=FALSE}

knitr::opts_chunk$set(cache=TRUE,
                      message=FALSE, warning=FALSE,
                      echo = FALSE,
                      fig.path='figs/',
                      cache.path = '_cache/',
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      })

library(dplyr)
library(ggplot2)
library(ggthemes)
library(ggbeeswarm)
library(showtext)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
font_add("Linux Libertine G", regular = "LinLibertine_R_G.ttf")

cal_folder <- 'D:/calibrations/phenofit/fagus_sylvatica/1000pres_1000abs/paper_data'

sim_folder <- 'D:/simulations/phenofit/backward/fagus_sylvatica/paper_data'

```

\hfill \break

## Fagus sylvatica leaf unfolding submodel

This model, called UniChill [@Chuine2000], is a sequential two-phase model (endodormancy and ecodormancy phases).  
The endodormancy phase begins at day $t_0$. 
The daily rate of chilling $R_c$ is defined as a threshold function of the daily mean temperature $T_d$:
$$ R_c(T_d) = \left\{
\begin{array}{ll}
      0 & T_d \geq T_b \\
      1 & T_d < T_b \\
\end{array} 
\right. $$
where $T_b$ is the threshold temperature below which the bud  accumulates chilling units.  
The endodormancy releases at day $t_c$ when the accumulated rate of chilling has reached the level $C_{crit}$:
$$\sum\limits_{t_0}^{t_c} R_c(T_d) \geq C_{crit} $$
Then, the ecodormancy phase begins. The daily rate of forcing $R_f$ is defined as a sigmoid function of the daily mean temperature $T_d$:
$$R_f(T_d) = \frac{1}{1 + e^{-d_T(T_d-T_{50})}} $$
where $d_T$ is the slope and $T_{50}$ the mid-response temperature.
Bud break occurs at day $t_f$ when the accumulated rate of forcing has reached the level $F_{crit}$:
$$\sum\limits_{t_c}^{t_f} R_f(T_d) \geq F_{crit} $$
Thus, the UniChill model has 6 parameters: $t_0$, $T_b$ and $C_{crit}$ for the first phase, $d_T$, $T_{50}$ and $F_{crit}$ for the second phase. 


```{r leafpardensity, include=TRUE, fig.height = 4, fig.width = 4, fig.align = "center", fig.cap = "Beech leaf unfolding model parameter density. Y-axis and X-axis limits are lower and upper bounds used during calibration.", fig.pos="H", fig.showtext=TRUE, eval=T}

par_names <- unique(data_leaf_par$var) 
names(par_names) <- c("t_0", "T_b", "d_T", "T_{50}", "C_{crit}", "F_{crit}")

parameter_name1 <- "T_b"
parameter_name2 <- "C_{crit}"
plot_name <- "plot_Vb_Ccrit"
breaks1 <- c(0, 5)
breaks2 <- c(50, 100)
source(file.path(wd, "scripts", "generate_density_plot.R"))

parameter_name1 <- "d_T"
parameter_name2 <- "F_{crit}"
plot_name <- "plot_d_Fcrit"
breaks1 <- c(-1.5, -0.5)
breaks2 <- c(40, 80)
source(file.path(wd, "scripts", "generate_density_plot.R"))

parameter_name1 <- "T_{50}"
parameter_name2 <- "F_{crit}"
plot_name <- "plot_e_Fcrit"
breaks1 <- c(10, 15)
source(file.path(wd, "scripts", "generate_density_plot.R"))

parameter_name1 <- "F_{crit}"
parameter_name2 <- "C_{crit}"
plot_name <- "plot_Fcrit_Ccrit"
breaks1 <- c(40, 80)
breaks2 <- c(50, 100)
source(file.path(wd, "scripts", "generate_density_plot.R"))
  
p <- plot_grid(plot_Vb_Ccrit + theme(legend.position = "none"), plot_Fcrit_Ccrit + theme(legend.position = "none"),
               plot_d_Fcrit + theme(legend.position = "none"), plot_e_Fcrit + theme(legend.position = "none"), 
          labels=c("(A)", "(B)", "(C)", "(D)"), ncol = 2, nrow = 2,
          hjust = 0, vjust = 1, align = "h", scale = 0.95)

legend <- get_legend(plot_Vb_Ccrit)

plot_grid(p, legend, ncol = 1, nrow = 2, rel_heights= c(0.9, 0.25))

```

```{r parametersplots, include=TRUE, fig.height = 4, fig.width = 6, fig.align = "center", fig.cap = "Impacts of replications of CMA-ES calibration, with PHENOFIT model and beech.", fig.pos="H", fig.showtext=TRUE, eval= F}

data_fruit_par <- species_parameters %>%
  dplyr::filter(grepl("fruit", species_parameters$var))
data_fruit_par$var_name <- factor(data_fruit_par$var, labels = c('fruitmat1' = parse(text=TeX('$aa$')),
                                                               'fruitmat2' = parse(text=TeX('$bb$')),
                                                               'fruitmat3' = parse(text=TeX('$F_{crit}$')),
                                                               'fruitmat4' = parse(text=TeX('$T_{opt}$')),
                                                               'fruitmat5' = parse(text=TeX('$Mat_{moy}$')),
                                                               'fruitmat6' = parse(text=TeX('$\\sigma$')),
                                                               'fruitmat7' = parse(text=TeX('$PFE_{50}$'))))

data_fruit_par$var_name <- factor(data_fruit_par$var, labels = c('fruitmat1' = parse(text=TeX('$aa$')),
                                                               'fruitmat2' = parse(text=TeX('$bb$')),
                                                               'fruitmat3' = parse(text=TeX('$cc$')),
                                                               'fruitmat4' = parse(text=TeX('$dd$')),
                                                               'fruitmat5' = parse(text=TeX('$ee$')),
                                                               'fruitmat6' = parse(text=TeX('$sig$')),
                                                               'fruitmat7' = parse(text=TeX('$gg$'))))

fruit_parameter_plot <- data_fruit_par %>%
  ggplot(aes(y = value, x = factor(1))) +
  geom_violin(fill = 'lightgrey', alpha= 0.2) +
  geom_hline(aes(yintercept = init), linetype = 'dashed', col = 'darkred') +
  geom_beeswarm(col = 'darkred', 
                alpha = 0.5, cex = 2.5, size = 0.4) +
  facet_wrap("var_name", scales="free",
             labeller = label_parsed) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(family = "Linux Libertine G"),
        axis.title.y = element_text(family = "Linux Libertine G"),
        strip.text.x = element_text(family = "Linux Libertine G", face = "bold", size = 11),
        axis.ticks.x=element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position="none") +
  geom_blank(aes(y = lb)) +
  geom_blank(aes(y = ub)) +
  labs(y = "")

data_frost_par <- species_parameters %>%
  dplyr::filter(grepl("frost", species_parameters$var))
data_frost_par$var_name <- factor(data_frost_par$var, labels = c('frost1' = parse(text=TeX('$FH_{max1}^{fr}$')),
                                                               'frost2' = parse(text=TeX('$FH_{max2}^{fr}$')),
                                                               'frost3' = parse(text=TeX('$FH_{min}^{fe}$')),
                                                               'frost4' = parse(text=TeX('$FH_{min}^{f1}$')),
                                                               'frost5' = parse(text=TeX('$Te_1$')),
                                                               'frost6' = parse(text=TeX('$Te_2$')),
                                                               'frost7' = parse(text=TeX('$FH_{max}^{tfe}$')),
                                                               'frost8' = parse(text=TeX('$FH_{max}^{tf1}$')),
                                                               'frost9' = parse(text=TeX('$FH_{max}^{pfe}$')),
                                                               'frost10' = parse(text=TeX('$FH_{max}^{pf1}$')),
                                                               'frost11' = parse(text=TeX('$NL_1$')),
                                                               'frost12' = parse(text=TeX('$NL_2$'))))
frost_parameter_plot <- data_frost_par %>%
  ggplot(aes(y = value, x = factor(1))) +
  geom_violin(fill = 'lightgrey', alpha= 0.2) +
  geom_hline(aes(yintercept = init), linetype = 'dashed', col = 'darkred') +
  geom_beeswarm(col = 'darkred', 
                alpha = 0.5, cex = 2.5, size = 0.4) +
  facet_wrap("var_name", scales="free",
             labeller = label_parsed) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(family = "Linux Libertine G"),
        axis.title.y = element_text(family = "Linux Libertine G"),
        strip.text.x = element_text(family = "Linux Libertine G", face = "bold", size = 11),
        axis.ticks.x=element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position="none") +
  geom_blank(aes(y = lb)) +
  geom_blank(aes(y = ub)) +
  labs(y = "")

data_senes_par <- species_parameters %>%
  dplyr::filter(grepl("senes", species_parameters$var))
data_senes_par$var_name <- factor(data_senes_par$var, labels = c('senes1' = parse(text=TeX('$P_b$')),
                                                               'senes2' = parse(text=TeX('$T_b$')),
                                                               'senes3' = parse(text=TeX('$\\alpha$')),
                                                               'senes4' = parse(text=TeX('$\\beta$')),
                                                               'senes5' = parse(text=TeX('$S_{crit}$')),
                                                               'senes6' = parse(text=TeX('$\\sigma$'))))
senes_parameter_plot <- data_senes_par %>%
  ggplot(aes(y = value, x = factor(1))) +
  geom_violin(fill = 'lightgrey', alpha= 0.2) +
  geom_hline(aes(yintercept = init), linetype = 'dashed', col = 'darkred') +
  geom_beeswarm(col = 'darkred', 
                alpha = 0.5, cex = 2.5, size = 0.4) +
  facet_wrap("var_name", scales="free",
             labeller = label_parsed) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(family = "Linux Libertine G"),
        axis.title.y = element_text(family = "Linux Libertine G"),
        strip.text.x = element_text(family = "Linux Libertine G", face = "bold", size = 11),
        axis.ticks.x=element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position="none") +
  geom_blank(aes(y = lb)) +
  geom_blank(aes(y = ub)) +
  labs(y = "")

plot_grid(fruit_parameter_plot, fruit_parameter_plot, labels=c("(A)", "(B)"), ncol = 2, nrow = 2,
          hjust = 0, vjust = 1, align = "h", label_size = 12, 
          label_colour = "#495057")



plot_grid(fruit_parameter_plot, mad_map, labels=c("(A)", "(B)"), ncol = 2, nrow = 1,
          hjust = 0, vjust = 1, align = "h", label_size = 12, rel_widths = c(1, 0.9),
          label_colour = "#495057")

```

```{r consensusfitnessmap, include=TRUE, fig.height = 4, fig.width = 7, fig.align = "center", fig.cap = "Median absolute deviation of beech \\textbf{(A)} leaf unfolding date and \\textbf{(B)} fitness, predicted with 100 calibrated parameter sets of PHENOFIT.", fig.pos="H", fig.showtext=TRUE}

mad_unfolding <- ggplot(data=unfold_date_data, aes(x = lon, y = unfold_date_data$lat)) + 
  geom_tile(aes(fill = mad), color = NA) +
  theme_void() +
  scale_fill_gradient2(name = stringr::str_wrap("Median absolute deviation", width = 30), 
                       low = "#0A9396", mid ="#e9d8a6", high = "#bb3e03", 
                       limits = c(0,max(unfold_date_data$mad)), 
                       midpoint = 60,
                       guide = guide_colourbar(title.position = "top", frame.colour = "black", frame.linewidth = 0.5, ticks = FALSE)) +
  theme(legend.title.align = 0.5, legend.direction = "horizontal", legend.position = "bottom", 
        legend.key.height = unit(0.2, 'cm'), legend.key.width = unit(1, 'cm'),
        legend.title = element_text(family = "Linux Libertine G", size = 9),
        legend.text = element_text(family = "Linux Libertine G"))

mad_fitness <- ggplot(data=fitness_data, aes(x = lon, y = lat)) + 
  geom_tile(aes(fill = mad), color = NA) +
  theme_void() +
  ylab("") +
  xlab("") +
  scale_fill_gradient2(name = stringr::str_wrap("Median absolute deviation", width = 30), 
                       low = "#0A9396", mid ="#e9d8a6", high = "#bb3e03", 
                       limits = c(0,max(fitness_data$mad)), 
                       midpoint = 0.25,
                       guide = guide_colourbar(title.position = "top", frame.colour = "black", frame.linewidth = 0.5, ticks = FALSE)) +
  theme(legend.title.align = 0.5, legend.position = "bottom", legend.direction = "horizontal",
        legend.key.height = unit(0.2, 'cm'), legend.key.width = unit(1, 'cm'),
        legend.title = element_text(family = "Linux Libertine G", size = 9),
        legend.text = element_text(family = "Linux Libertine G"))

plot_grid(mad_unfolding, mad_fitness, labels=c("(A)", "(B)"), ncol = 2, nrow = 1,
          hjust = 0, vjust = 1, align = "h", label_size = 12,
          label_colour = "#495057")

fitness_data_fsyl <- left_join(fsyl_presabs, fitness_data)

```

The median standard deviation of unfolding date across Europe was about `r round(median(unfold_date_data$sd),1)` days. On beech presence points, it was about `r round(median(unfold_date_data_fsyl[unfold_date_data_fsyl$pres ==1, 'sd']),1)` days. Nearly `r round(nrow(unfold_date_data[unfold_date_data$mad < 30,])/nrow(unfold_date_data)*100,1)`% of cells had a median absolute deviation lower than 30 days (\hyperref[fig:consensusfitnessmap]{Figure F.2.A.}).
The median standard deviation of fitness across Europe was about `r round(median(fitness_data$sd),3)`. On beech presence points, it was about `r round(median(fitness_data_fsyl[fitness_data_fsyl$pres ==1, 'sd']),3)`. Nearly `r round(nrow(fitness_data[fitness_data$mad < 0.1,])/nrow(fitness_data)*100,1)`% and `r round(nrow(fitness_data[fitness_data$mad < 0.2,])/nrow(fitness_data)*100,1)`% of total cells had a median absolute deviation lower than 0.1 and 0.2 respectively (\hyperref[fig:consensusfitnessmap]{Figure F.2.B.}).

```{r unfoldingdatesbackforw, include=TRUE, fig.height = 4, fig.width = 7, fig.align = "center", fig.cap = "Mean leaf unfolding day of beech with \\textbf{(A)} best CMA-ES calibrated parameters and \\textbf{(B)} expert parameters. Values above June solstice day (167) are in grey. Note that PHENOFIT model assign a value of 365 when unfolding has not happened at all due to climate conditions.", fig.pos="H", fig.showtext=TRUE}

best_unfoldingdate <- read_mean_outputvalue_phenofit(file.path(file.path(sim_folder, "CMAES"), phenofit_fagus_best_rep, "LeafUnfoldingDate.txt"))
backward_unfoldingdate <- data.frame(lat= alt$lat, lon =alt$lon, date = t(best_unfoldingdate))
backward_unfoldingdate[backward_unfoldingdate$date > 172, "date"] <- NA # 21 june

backward_unfolding_map <- ggplot(data=backward_unfoldingdate, aes(x = lon, y = lat)) + 
  geom_tile(aes(fill = date), color = NA) +
  theme_void() +
  scale_fill_gradient2(name = stringr::str_wrap("Leaf unfolding day", width = 30), 
                       low = "#0A9396", mid ="#e9d8a6", high = "#bb3e03", 
                       limits = c(100,max(backward_unfoldingdate$date)), 
                       midpoint = 140,
                       guide = guide_colourbar(title.position = "top", frame.colour = "black", frame.linewidth = 0.5, ticks = FALSE),
                       na.value = "grey50") +
  theme(legend.title.align = 0.5, legend.direction = "horizontal", legend.position = "bottom", 
        legend.key.height = unit(0.2, 'cm'), legend.key.width = unit(1, 'cm'),
        legend.title = element_text(family = "Linux Libertine G", size = 9),
        legend.text = element_text(family = "Linux Libertine G"))


forward_sim_folder <- "D:/simulations/phenofit/forward"
unfoldingdate <- read_mean_outputvalue_phenofit(file.path(forward_sim_folder, "fagus_sylvatica/EvolLett2019", "LeafUnfoldingDate.txt"))
forward_unfoldingdate <- data.frame(lat= alt$lat, lon =alt$lon, date = t(unfoldingdate))
forward_unfoldingdate[forward_unfoldingdate$date > 172, "date"] <- NA

forward_unfolding_map <- ggplot(data=forward_unfoldingdate, aes(x = lon, y = lat)) + 
  geom_tile(aes(fill = date), color = NA) +
  theme_void() +
  scale_fill_gradient2(name = stringr::str_wrap("Leaf unfolding day", width = 30), 
                       low = "#0A9396", mid ="#e9d8a6", high = "#bb3e03", 
                       limits = c(100,max(forward_unfoldingdate$date)), 
                       midpoint = 140,
                       guide = guide_colourbar(title.position = "top", frame.colour = "black", frame.linewidth = 0.5, ticks = FALSE),
                       na.value = "grey50") +
  theme(legend.title.align = 0.5, legend.direction = "horizontal", legend.position = "bottom", 
        legend.key.height = unit(0.2, 'cm'), legend.key.width = unit(1, 'cm'),
        legend.title = element_text(family = "Linux Libertine G", size = 9),
        legend.text = element_text(family = "Linux Libertine G"))


plot_grid(backward_unfolding_map , forward_unfolding_map, labels=c("(A)", "(B)"), ncol = 2, nrow = 1,
          hjust = 0, vjust = 1, align = "h", label_size = 12,
          label_colour = "#495057")

fitness_data_fsyl <- left_join(fsyl_presabs, fitness_data)

```


